Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-20 manuscript.tex
more cuts

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data. Could body position be successfully recorded over the desired period? If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods. Based on past results \autocite{Survey}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing). Caregivers indicated in advance which size would fit their infant, and to ``size down'' if on the cusp to ensure a snug fit and reduce sensor movement. Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.

Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors. However, LENA® data were not analyzed as part of the current study.

@@

Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Figure \ref{fig:exemplar-timeline} shows an exemplar timeline of the entire procedure and recording periods for a single participant. On the day of the visit, a researcher arrived at the participant's home in the morning and prepared all the equipment at the doorstep. In order to synchronize all three recording devices, the researcher began by first turning on the video camera and the LENA audio device (the IMUs were already configured and placed before arrival because they required a proprietary sensor dock to begin recording in the laboratory). To create an easily recognizable synchronization point between the video recording and IMU data, the researcher dropped or struck the leggings (containing the IMUs) on a surface in view of the camera, as in \textcite{FranchakScott2021}. All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

Afterwards, the researcher asked the caregiver to complete a number of guided activities with their infant. Within view of the camera, the caregiver was asked to place their infant in several different positions: lying supine, lying prone, sitting on the floor, standing upright, held by the caregiver while the caregiver walked back and forth, crawling, walking, and sitting in a restrained seat (e.g., high chair). Depending on the infants' age and motor skill level, the positions could be done independently or were completed with assistance from the caregiver. The researcher kept time to ensure at least 1 minute of behavior for each activity. Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional training data with the infant in positions that would be typical of play.

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, diaper changes, and trips out of the house. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline}) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline}) so that those times could be excluded from analysis. The following day a researcher picked up the equipment, verified that the paperwork was signed, and compensated the participant.

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was automatically recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

Supine was coded when the infant was lying on their back or was reclined up to a 45 degree angle. Supine was also coded in the rare cases when the infant was laying on their side. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). When an infant was carried by a caregiver, held was coded. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the sensor garment was adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}. The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.

The resulting windowed dataset was used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models \autocite{Breiman2001} were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package \autocite{LiawWiener2002}. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R \autocite{tidymodels} was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550; however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides, to our knowledge, the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

The first set of analyses use data from the proximal comparison to determine the ``best case'' accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work \autocite{FranchakScott2021}, we compared two types of models: \emph{group models} and \emph{individual models}. To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75\% of a recording session, then tested against the (temporally) last 25\% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92). Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior. Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy. Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93). Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was far lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

Similar to past work \autocite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used. Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9). The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819). Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and thigh has group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors. Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.

@@

Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced. For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82). Guidelines for interpreting kappa statistics \autocite{LandisKoch1977} consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.

>>>>

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be scored as sitting. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data. Could infants wear the sensor garment throughout the desired period? If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position. Based on past results \autocite{Survey}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration). A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing). Caregivers indicated in advance which size would fit their infant, and to ``size down'' if between sizes to ensure a snug fit and minimize sensor movement. Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.

Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours. We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors (LENA® data were not analyzed in the current study).

@@

Videos were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the same room as the infant. Although recordings lasted 3 hours, they were divided into two video files of approximately 90 minutes temporally separated by a gap of 40-45 s. Caregivers received a log sheet to record times that infants napped and times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Figure \ref{fig:exemplar-timeline} shows an exemplar timeline of the entire procedure and recording periods for a single participant. A researcher arrived at the participant's home in the morning and set each device to record while at the doorstep. To create a recognizable synchronization point between the video recording and IMU data, the researcher dropped the sensor garment containing the IMUs on a surface in view of the camera, as in \textcite{FranchakScott2021}. All the equipment---once recording and with synchronization information recorded---was placed inside a large bucket and left outside the family's front door. The researcher then called the caregiver on the phone and walked them through a set of procedures needed to properly set up the equipment and record video for ground truth human annotation of body position. At the start of this ``guided call'', the caregiver was instructed to place the camera in an area that captured the majority of the room. Next, they were asked to put the pair of leggings and shirt on their infant, with the researcher providing guidance about how to correctly orient the garments.

Afterwards, the researcher asked the caregiver to complete a number of guided activities with their infant. Within view of the camera, the caregiver was asked to place their infant in several different positions: lying supine, lying prone, sitting on the floor, standing upright, held by the caregiver while the caregiver walked back and forth, crawling, walking, and sitting in a restrained seat (e.g., high chair). Depending on the infants' age and motor skill level, the positions could be done independently or were completed with assistance from the caregiver. The researcher kept time to ensure at least 1 minute of behavior for each activity. Once completed, the caregiver was then instructed to play with their infant for 10 minutes within view of the camera to collect additional ground truth data.

Afterwards, they were asked to go about their day as usual with the infant wearing the sensor garment until their bedtime, only taking off the sensor for naps, baths, diaper changes, and trips out of the house. The caregiver logged the times the sensors were removed (blank areas in the timeline in Figure \ref{fig:exemplar-timeline}) or the child took a nap (gray areas in the timeline in Figure \ref{fig:exemplar-timeline}) so that those times could be excluded from analysis. The following day a researcher picked up the bucket of equipment.

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). The second video file comprised the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface). Held was coded when infants were carried off of the groun. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the sensor garment was adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Data were reduced in time by creating overlapping moving windows (4-s long, comprising 250 samples) starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}. For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.

The resulting windowed dataset was used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set (further described below). Random forest models \autocite{Breiman2001} used the training set to learn the body position label for each window from the set of 436 motion features using the \emph{randomForest} package \autocite{LiawWiener2002}. The resulting random forest model could later be applied to a set of testing data with the \emph{predict} function.

@@

We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording.

@@

The first set of analyses use data from the proximal comparison to determine the ``best case'' accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work \autocite{FranchakScott2021}, we compared two types of models: \emph{group models} and \emph{individual models}. To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75\% of a recording session, then tested against the (temporally) last 25\% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92). Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies \autocite{FranchakScott2021,AiraksinenRasanen2020}, both models approached the level of agreement found between two human coders (\emph{M} = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior. Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy. Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93).

As in \textcite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used. Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9). The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819). Accuracy for some pairs approached 4-sensor accuracy: Left ankle and thigh had group model accuracy of \emph{M} = NA and individual model accuracy of \emph{M} = NA, only 2.2\% and 1.5\% worse than using all four sensors. Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.

@@

Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and we provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82). Guidelines for interpreting kappa statistics \autocite{LandisKoch1977} consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
